A game-theory modeling approach to utility and strength of interactions dynamics in biomedical research social networks

### 2017-04-02 casm.tex
One problematic reference fixed and changed of position the conclusions all the way to the last paragraph.
proof-read to just before methodology

{\color{red}The analysis of cooperation in scientific research has been the
subject of a number of studies
\cite{VermeulenPenders:2013,Newman:2001,Newman:2004,Elango-etal:2012,HernandezLemus:2013,Strasser:2006,Strasser:2012}. This
is not surprising since cooperation and competition are quite
important in today's academic success. Moreover, the quantitative
analysis of cooperation in science also has a long history of more
than 50 years
\cite{Mali-etal:2012,Scharnhorst-etal:2012}. Particularly appealing to
researchers studying this topic is the use of network analysis,
especially in the area of bibliometrics and scientometrics such as in
citation networks that date as back as 1972 \cite{Garfield:1972},
co-authorship and collaboration networks also have a long tradition
\cite{Todorov-etal:1991,Wagner-etal:2017}.}\\

{\color{red}The approach we follow here articulates game theory together with
complex network analysis. Putting together game-theoretic modeling on
complex networks is common in studies about the emergence of
cooperation
\cite{SzaboFath:2007,Nowak-etal:92,OshtukiNowak:2006,Santos-etal:2005,Santos-etal:2006}.
However, in the context of science studies, it is far from being a
unified, active field of research. There is no common set of questions
and problems or a coherent body of models and theory aimed to
understand the quirks and twists of cooperation in scientific
communities. It is more likely to say that work of this kind is
scarce and their implementatoin is on different situations related to
the social life of science, where sometimes models are run on science
collaboration networks because access to these kind of data is
relatively easy to access.}\\

{\color{red}Nevertheless, there is work done. For instance Wardil and Hauert have
analyzed how does collaboration happen within co-authorship networks
\cite{Wardil-etal:2015}. Using the \emph{Physical Review} journal
historic database, these authors show the payoff differences between
being the first author and the last author, where the first author is
a cooperator that takes the highest load of work, while the last
author is the \emph{laird} that can benefit from the work of the
former and get engaged in many other projects. They argue that
cooperation in a multi-authored paper resambles more to a snow-drift
game than a prisoner's dilemma situation, since when co-authoring a
paper, every body wins but each author pays a different price. Though
this work takes into account the topology of the network as
informative regarding cooperation patterns, the role of topology
framing or fostering cooperation is not part of the study.} \\

{\color{red}Game theory over complex scientific information and collaboration
networks has also attracted attention. Hanauske has applied
\emph{Evolutionary Game Theory} to have a better understanding of the
market of scientific information, that involves scientists as
producers but also journal, that are in charge of its
dissemination. One of his models connects scientist interest in
maximazing their reputation and their decissions to publish in an Open
Access Journal or in a journal that works under the traditional
publication framework. He has also modeled a game-theoretic
interactions between scientists and journals for both players to go
open access or not. Only this second model incorporates networks
analysis \cite{Hanauske-etal:2012}.}\\

{\color{red}Another interesting behavior explored by game-theoretical modeling and
complex networks has been done by Li and Cao \cite{Li-etal:2009}. This
authors analyze the role of punishment in a science collaboration
network. More precisely, they implemented an ultimatum bargaining
model to understand the impact of the topology of a social network and
the emergence of a population of punishers. In the ultimatum game,
there are two players, one of them (the proponent) suggest to split a
sum in two and its connected neighbor (the respondent), can decide to
accept or reject the offer as it may be perceived more or less
fair. If the respondent accepts, the sum is divided according to the
offer. If the offer is rejected both players do not get
anything. According to the authors, the act of rejecting a selfish
offer can be interpreted as a punishment since the player that rejects
it is paying a cost of not receiving anything but the proponent looses
even more, since his offer was selfish. Their results show that the
population of punishers will emerge as a phase transition as the
degree of selfishness increses among players. Moreover, they found a
dependence relation between the tolerance to selfish behaviors and the
topologies of networks.}\\

>>>>

{\color{red}The analysis of cooperation in scientific research has been the
subject of a number of studies
\cite{VermeulenPenders:2013,Newman:2001,Newman:2004,Elango-etal:2012,HernandezLemus:2013,Strasser:2006,Strasser:2012}. This
is not surprising since cooperation and competition are quite
important in today's academic success. Moreover, the quantitative
analysis of cooperation in science also has a long history of more
than 50 years
\cite{Mali-etal:2012,Scharnhorst-etal:2012}. Particularly appealing to
researchers studying this topic is the use of network analysis,
especially in the area of bibliometrics and scientometrics such as in
citation networks that date as far back as 1972 \cite{Garfield:1972},
co-authorship and collaboration networks also have a long tradition
\cite{Todorov-etal:1991,Wagner-etal:2017}.}\\

{\color{red}The approach we follow here articulates game theory together with
complex network analysis. Putting together game-theoretic modeling on
complex networks is common in studies about the emergence of
cooperation
\cite{SzaboFath:2007,Nowak-etal:92,OshtukiNowak:2006,Santos-etal:2005,Santos-etal:2006}.
However, in the context of science studies, it is far from being a
unified, active field of research. There is no common set of questions
and problems or a coherent body of models and theory aimed to
understand the quirks and twists of cooperation in scientific
communities. It is more earnest to say that work of this kind is
rather scarce and focused around situations related to
the social life of science, where sometimes models are run on
collaboration networks because access to that kind of data is
relatively easy.}\\

{\color{red}Nevertheless, there is work done. For instance Wardil and Hauert have
analyzed how collaboration happens within co-authorship networks
\cite{Wardil-etal:2015}. Using the \emph{Physical Review} journal
historic database, these authors show the payoff differences between
being the first author and the last author, where the first author is
a cooperator that takes the highest load of work, while the last
author is the \emph{laird} that can benefit from the work of the
former and get engaged in many other projects. They argue that
cooperation in a multi-authored paper is a game more similar to ``snow-drift''
than to ``prisoner's dilemma'' since when co-authoring a
paper everybody wins but each author pays a different price. Though
this work takes into account the topology of the network as
informative regarding cooperation patterns, the role of topology
framing or fostering cooperation is not part of the study.} \\

{\color{red}

Game theory over complex scientific information and collaboration
networks has also attracted attention. To have a better understanding
of the market of scientific information Hanauske has applied
\emph{Evolutionary Game Theory} involving scientists as producers but
also journals as disseminators. One of his models connects scientists'
interest in maximazing their reputation and their decissions to
publish in an Open Access Journal or in a journal that works under the
traditional publication framework. He has also modeled game-theoretic
interactions between scientists and journals concerning the decision
of both players to go open access or not. Only this second model
incorporates networks analysis \cite{Hanauske-etal:2012}.}\\

{\color{red}

Another interesting behavior explored by game-theoretical modeling and
complex networks has been done by Li and Cao \cite{Li-etal:2009}. This
authors analyze the role of punishment in a science collaboration
network. More precisely, they implemented an ultimatum bargaining
model to understand the impact of the topology of a social network and
the emergence of a population of punishers. In the ultimatum game,
there are two players, one of them (the proponent) suggests to split a
sum in two and its connected neighbor (the respondent) can decide to
accept or reject the offer as it may be perceived more or less
fair. If the respondent accepts, the sum is divided according to the
offer. If the offer is rejected both players get nothing. According to
the authors, the act of rejecting a selfish offer can be interpreted
as a punishment since the player that rejects it is paying the cost of
not receiving anything, but because of a selfish offer the proponent
looses even more. Their results show that a phase transition will
emerge in the population of punishers as the degree of selfishness
increses among players. Moreover, they found a dependence relation
between the tolerance to selfish behaviors and the topologies of
networks.}\\
